### 2022-02-03 analysis/paper/paper.Rmd
latest edits from TB

---
title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
keywords: |
  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
---

Keywords: `r rmarkdown::metadata$keywords`

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

```

```{r load-libraries}
library(tidyverse)
library(magrittr)
library(janitor)
library(here)
library(lubridate)
library(FSA)
library(english)
library(lubridate)
library(tidytable)
library(kableExtra)
library(lemon)
library(gridExtra)

# format NA values in tables
options(knitr.kable.NA = '-')

# set default theme for figures
#theme_set(theme_bw())
theme_set(theme_classic())

# when knitting to Word, use this
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

if(doc.type == 'docx') {
  options(knitr.table.format = "pandoc")
}

```

\newpage

## Abstract

Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery efforts in the Columbia River basin. Reservoirs and tailraces associated with dams on the Snake and Columbia river have created favorable conditions for fishes known to consume juvenile salmonids. The native Northern Pikeminnow are one of these fishes and have been estimated to consume ~8% of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually. Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above these reservoir systems. In this study, we examine the abundance of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho, known as Deadwater Slough, and consider potential impacts of them on local Chinook Salmon populations. The reach was (at least partially) formed by the failure of a small mining dam in a nearby drainage during the late 19th century and contains historically important overwinter rearing habitat and is a migration corridor for Chinook Salmon as well as other Endangered Species Act-listed species including steelhead and Sockeye Salmon. We estimated the abundance of Northern Pikeminnow in Deadwater Slough to be 27,874 in the fall and 19,499 in the spring during the peak emigrations of juvenile Chinook Salmon. Assuming a scenario where 60% of the Northern Pikeminnow diet is fish, of which 50% is Chinook Salmon, we estimated the Northern Pikeminnow population to consume 95,120 juvenile Chinook Salmon, annually. In the most conservative scenario where 30% of diet is fish and 30% of fish are juvenile Chinook Salmon, 47,079 juveniles are consumed. Using a sensitivity analysis, we further considered potential impacts to adult returns to local populations. Given the size of the Salmon River, the estimated Northern Pikeminnow densities for Deadwater Slough are remarkable and predation from the population likely has consequential impacts on Chinook Salmon in the Upper Salmon River major population group, thereby hindering local recovery efforts.

\newpage

# Introduction

The Snake and Columbia rivers drain a network of inland tributaries that provide essential spawning and rearing habitat for anadromous Pacific salmon *Oncorhynchus* spp. and steelhead *O. mykiss*. These fishes were historically abundant throughout the basin and consequently have tremendous ecological, cultural, and economic value [@Atlas2021; @Lewis2019]. Over the last century, salmonid stocks throughout the Snake and Columbia rivers have undergone significant declines related to habitat modification in the freshwater environment (e.g., removal of beavers, river channel simplification, water withdrawals, hydrosystem development) and poor marine conditions, affecting their survival and recruitment at multiple life stages [@Crozier2020; @Clark2020; @Justice2017]. Consequently, action agencies have made considerable investments in the rehabilitation of tributary ecosystems and evaluation of factors attributed to the species decline [@White2021; @Roni2018].

One potentially important, but perhaps under-appreciated source of salmonid mortality is predation on emigrating juveniles by piscivorous fishes. In the lower mainstem Snake and Columbia rivers it is estimated that predation on out-migrating salmonids during peak emigration has a significant negative impact on the overall population and success of recovery [@Fresh2003; @Winther2020]. Dams and reservoirs in the Columbia River are the primary locations associated with high rates of piscine predation on salmonids [@Petersen1994; @Ward1995]. There are generally two mechanisms that explain these high predation zones. First, migration (movement) rates of juvenile salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation. Second, reservoirs and downstream tailraces associated with dams on the Snake and Columbia rivers create favorable slow-water habitat for predatory fishes known to consume juvenile salmonids, such as the Northern Pikeminnow *Ptychocheilus oregonsis*. Northern Pikeminnow tolerate and thrive in relatively warm, slow water habitats [@Wydoski2003] and consequently have benefited from dams on the Columbia River, becoming abundant predators of salmonid outmigrants [@Knutsen1999]. Indeed, Northern Pikeminnow are estimated to consume ~8% (16.4 million) of the approximately 200 million juvenile salmonids emigrating through the lower Snake and Columbia Rivers, annually [@Beamesderfer1996].

```{r deadwater-attr}
# set some deadwater attributes
ddw_rkms = 1.871               # length km
ddw_m2 = 119431                # area square m
ddw_avg_width_m = 68           # measured avg width
ddw_acres = ddw_m2 / 4047
ddw_hectares = ddw_m2 / 10000

```

Most predation studies have focused on reservoirs [@Murphy2021] and mainstem reaches [@Zimmerman1999; @Tabor1993; @Ward1995; @Shively1996] of the Columbia River basin whereas considerably less is known about the interaction between upstream, river-dwelling piscivorous fishes and their impacts on salmonid recovery above the reservoir systems [@Rubenson2020]. Upstream habitats containing slower water velocities and other attributes that support piscivorous predators may overlap with essential habitat for some salmonid species and life stages. One such example is Deadwater Slough, an approximately `r round(ddw_rkms, 1)` km long reach of unnaturally slow and deep water in the Salmon River, Idaho. The Deadwater Slough is in a river reach containing historically important overwinter rearing habitat, includes the emigration corridor for juvenile Chinook Salmon *O. tshawytscha*, and is inhabited by piscivorous predators, including Northern Pikeminnow and Smallmouth Bass *Micropterus dolomieu*. Further, the slough is part of the migratory pathway of the endangered Snake River Sockeye Salmon *O. nerka* population [@Axel2015] and several populations of threatened Snake River steelhead.

Quantifying mortality associated with piscine predation in upstream habitats of the Salmon River is of particular interest for spring/summer-run Chinook Salmon (hereafter Chinook Salmon) in the Upper Salmon major population group (MPG). The Salmon River was historically the most productive tributary for Chinook Salmon in the Columbia River basin [@Nemeth1999], but those populations have become depleted in recent decades. Currently, the Upper Salmon MPG supports eight independent, extant populations; this includes Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River populations [@NOAA2017]. The Lemhi River historically contained the largest Chinook Salmon population within the MPG and is therefore of particular interest for restoration actions [@Zimmerman2012]. The Upper Salmon MPG are stream-type Chinook Salmon and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014a]. The DSR migrants leave their natal area as subyearlings between June and November and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts. Alternatively, NRR migrants remain in their natal areas for approximately one year after emergence until emigration to the ocean as smolts. Diversity of migration tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency [@Dodson2013]. Deadwater Slough represents an important habitat for the Upper Salmon MPG as is supports rearing and migration of juveniles from all populations including fall DSR and spring NRR migrants.

Recent studies have examined the downstream movement, distribution, and apparent survival (hereafter 'survival') of juvenile salmonid emigrants through the Salmon River, including the Deadwater Slough reach. Sockeye Salmon migrating through the mainstem Salmon River during spring were estimated to have substantially reduced survival in the Deadwater Slough reach [@Axel2015]. Similarly, emigrating DSR Chinook Salmon had an approximate 10% reduction in transition probability through the Deadwater Slough compared to surrounding reaches during fall and early winter months [@Ackerman2018; @Porter2019]. Low survival was attributed to low-velocity water delaying movement rates, thereby increasing predation risk.

In this study, we estimated the abundance of Northern Pikeminnow in the Deadwater Slough and evaluated potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon. We hypothesize that high densities of piscivorous predators in the Deadwater Slough may explain the reduced survival observed for juvenile Chinook Salmon. To test this, our objectives were four-fold:

1. Estimate the abundance of Northern Pikeminnow in the Deadwater Slough during the peaks of fall DSR and spring NRR juvenile emigrations;
2. Document predation on juvenile Chinook Salmon during the DSR and NRR emigration periods using gastric lavage;
3. Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population on DSR and NRR Chinook Salmon emigrants at Deadwater Slough;
4. Estimate how consumption of juvenile Chinook Salmon emigrants at Deadwater Slough by Northern Pikeminnow may impact adult returns to the Upper Salmon MPG.

# Methods

## Study Site

The Deadwater Slough is an approximately `r round(ddw_rkms, 1)` km section of the mainstem Salmon River located roughly 6 river kilometers (rkm) downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)). The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River. Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event and upslope instability that deposited substantial amounts of sediment at the confluence of the Salmon River [@Emerson1973]. That sediment formed an alluvial fan which created a slow, deep section in the river, spanning approximately `r round(ddw_hectares)` hectares, averaging `r round(ddw_avg_width_m)` m wide, and resembling a small reservoir.

## Northern Pikeminnow Demographics

We estimated the population size of piscine predators in the Deadwater Slough using a mark-recapture survey design and a catch per unit effort (CPUE) approach. Predators were sampled near the peaks of the fall DSR and spring NRR emigrations. During our initial survey in 2019, Northern Pikeminnow were by far the most prevalent piscine predator in Deadwater Slough and consequently became our focal taxa. The initial intent was to survey during the emigration of the 2018 brood year (fall 2019 and spring 2020), however logistical constraints during the onset of the Covid-19 pandemic delayed the spring 2020 survey until fall 2020. Although this allowed an additional survey during the fall DSR emigration, surveys during the spring NRR emigration were still lacking. Consequently, an additional week of surveying was added in spring 2021. The fall surveys were constrained to two weeks to minimize Northern Pikeminnow immigration/emigration.

Multiple capture methods were employed during the initial fall 2019 survey to reduce selectivity and bias for species and size classes. Methods included raft electrofishing, fyke netting, snorkeling, and angling. After evaluating all methods, angling was the most effective method for capturing Northern Pikeminnow while also minimizing potential impacts to ESA-listed adult steelhead that were present in the Salmon River during our surveys. Ineffective methods were abandoned in subsequent years and the following analyses will focus on fish captured by angling, unless otherwise noted.

Our study relied heavily on volunteer anglers who were permitted to fish anywhere within Deadwater Slough. Each survey day, anglers boated or hiked their catch (periodically or upon filling a livewell) to a processing station at the boat ramp, approximately 500 m downstream from the top of the slough. For each fish, we recorded the date of capture, species, total length (TL; mm), and whether the fish was previously marked. Unmarked fish were given a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique to each day and then were released. Released fish were distributed throughout Deadwater Slough to facilitate mixing back into the population. Fish that died prior to release were included in the total capture (or recapture) count, but not included in the number of marked fish available for recapture after that date. Finally, we recorded the angling start and end time for each crew (person or combinations of persons) to calculate CPUE as the number of Northern Pikeminnow caught per angler hour.

```{r capture-data}
raw_df = readxl::read_excel(here("analysis/data/raw_data/Deadwater_predation_mark_recapture_11.09.21.xlsx")) %>%
  clean_names() %>%
  mutate(across(date,
                ymd)) %>%
  mutate(across(c(start_time, end_time),
                ~ ymd_hms(paste(year(date), month(date), day(date), hour(.), minute(.), second(.))))) %>%
  # break days into "sampling events"
  mutate(event_name = if_else(month(date) > 8,
                              paste("Fall", year(date), sep = "_"),
                              paste("Spring", year(date), sep = "_")))

# capture/mark/recpature data frame; focus on pikeminnow, to be used for mark-recap models
cmr_df = raw_df %>%
  filter(count > 0) %>%                                     # filter out effort-only records
  filter(species == "Northern Pikeminnow") %>%              # focus on NPM
  filter(method == "Angling") %>%
  select(event_name, 
         date,
         method, 
         species, 
         count, 
         length, 
         fish_status, 
         capture_type) %>%
  # if fish_status = Live & capture_type = Mark, set as mark_release as TRUE
  # if capture_type = Recapture, set recap to TRUE
  mutate(mark_release = if_else(fish_status == "Live" &     
                                capture_type == "Mark",
                                T, F),
         recap = if_else(capture_type == "Recapture",
                         T, F))

```

```{r effort}
# What size (mm) is considered a large pikeminnow?
large_size = 380

# catch per unit effort
cpue_df = raw_df %>%
  select(event_name,
         site,
         date,
         crew,
         crew_count,
         method,
         equipment,
         start_time:total_effort,
         total_day_effort) %>%
  distinct() %>%
  filter(method == "Angling") %>%
  mutate(across(total_effort,
                ~ . * crew_count)) %>%
  rename(total_effort_min = total_effort) %>%
  group_by(event_name,
           date) %>%
  summarise(across(c(total_effort_min,
                     crew_count),
                   sum),
            .groups = "drop") %>%
    left_join(cmr_df %>%
              group_by(event_name, species, date) %>%
              summarise(n_fish = n(),
                        n_large_fish = sum(length >= large_size,
                                           na.rm = T),
                        .groups = "drop")) %>%
  mutate(catch_per_hr = n_fish / (total_effort_min / 60),
         catch_per_angler_day = n_fish / crew_count,
         catch_lg_per_angler_day = n_large_fish / crew_count)

```

```{r lincoln-petersen-estimator}
# using Lincoln-Petersen model, combining first 4 and last four days
# define mark and recapture periods
cap_periods = tibble(event_name = "Fall_2019",
                     cap_period = "Mark",
                     int = interval(start = ymd('20191112'),
                                    end = ymd('20191114'))) %>%
  bind_rows(tibble(event_name = "Fall_2019",
                   cap_period = "Recapture",
                   int = interval(start = ymd('20191119'),
                                  end = ymd('20191121')))) %>%
  bind_rows(tibble(event_name = "Fall_2020",
                   cap_period = "Mark",
                   int = interval(start = ymd('20201020'),
                                  end = ymd('20201023')))) %>%
  bind_rows(tibble(event_name = "Fall_2020",
                   cap_period = "Recapture",
                   int = interval(start = ymd('20201027'),
                                  end = ymd('20201030'))))

# add cap_period column, fill w/ NA
cmr_df$cap_period = NA_character_
# fill cap_period w/ Mark or Recapture
for(i in 1:nrow(cap_periods)) {
  my_int = cap_periods$int[i]
  index = which(cmr_df$date %within% my_int)
  cmr_df$cap_period[index] = cap_periods$cap_period[i]
}

# Lincoln-Peterson Model
LP_mod = cmr_df %>%
  filter(event_name != "Spring_2021") %>%
  group_by(event_name, species) %>%
  summarise(M = sum(cap_period == "Mark" & fish_status == "Live", na.rm = T),
            n = sum(cap_period == "Recapture"),
            m = sum(cap_period == "Recapture" & capture_type == "Recapture", na.rm = T),
            .groups = "drop") %>%
  nest(data = -c(event_name, species)) %>%
  crossing(model = c("Petersen",
                     "Chapman")) %>%
    mutate(mr_model = map2(data,
                       model,
                      .f = function(x, y) {
                        with(x,
                             FSA::mrClosed(M = M,
                                          n = n,
                                          m = m,
                                          method = y))
                        }),
         N_summ = map(mr_model,
                      .f = summary,
                      incl.SE = T),
         CI = map(mr_model,
                  .f = confint)) %>%
    mutate(N = map_dbl(N_summ,
                     .f = function(x) x[1]),
         SE = map_dbl(N_summ,
                      .f = function(x) x[2]),
         Lci = map_dbl(CI,
                         .f = function(x) x[1]),
         Uci = map_dbl(CI,
                         .f = function(x) x[2])) %>%
    select(-N_summ, -CI)

```

```{r schnabel-estimator, eval = T}
# using Schnabel and Schumacher-Eschmeyer estimators
N_mods = cmr_df %>%
  # filter out spring 2021 (no recapture)
  filter(event_name != "Spring_2021") %>%
  group_by(event_name, species, date) %>%
  summarise(n = n(),
            m = sum(capture_type == "Recapture", na.rm = T),
            u = sum(capture_type == "Mark" | is.na(capture_type)),
            R = sum(fish_status == "Live", na.rm = T),
            .groups = "drop") %>%
  group_by(event_name, species) %>%
  mutate(M = lag(R - m),
         M = replace_na(M, 0),
         M = cumsum(M)) %>%
  nest() %>%
  crossing(model = c("Schnabel",
                     "SchumacherEschmeyer")) %>%
  mutate(mr_model = map2(data,
                         model,
                         .f = function(x, y) {
                           with(x,
                                mrClosed(n = n,
                                         m = m,
                                         R = R,
                                         method = y,
                                         chapman.mod = TRUE))
                         }),
         N = map_dbl(mr_model,
                     .f = summary),
         CI = map(mr_model,
                  .f = confint,
                  conf.level = 0.95),
         Lci = map_dbl(CI,
                       .f = function(x) x[1]), 
         Uci = map_dbl(CI,
                       .f = function(x) x[2])) %>%
  ungroup() %>%
  select(-CI)

```

```{r assump-violation}
# assume that marked fish aren't available for recapture until 48 days after release
# using Schnabel estimator
N_mods_viol = N_mods %>%
  filter(model == "Schnabel") %>%
  select(event_name:model) %>%
  mutate(data = map(data,
                    .f = function(x) {
                      x %>%
                        mutate(R_lag = lag(R),
                               R_lag = replace_na(R_lag,
                                                  0))
                    })) %>%
  mutate(mr_model = map2(data,
                         model,
                         .f = function(x, y) {
                           with(x,
                                mrClosed(n = n,
                                         m = m,
                                         R = R_lag,
                                         method = y,
                                         chapman.mod = TRUE))
                         }),
         N = map_dbl(mr_model,
                     .f = summary),
         CI = map(mr_model,
                  .f = confint),
         Lci = map_dbl(CI,
                       .f = function(x) x[1]),
         Uci = map_dbl(CI,
                       .f = function(x) x[2])) %>%
  ungroup() %>%
  select(-CI) %>%
  mutate(model = "Schnabel - Delayed Mixing")
```

```{r abund-est}
all_est = LP_mod %>%
  select(event_name, species, model, N:Uci) %>%
  bind_rows(N_mods %>%
             select(event_name, species, model, N:Uci)) %>%
  bind_rows(N_mods_viol %>%
             select(event_name, species, model, N:Uci)) %>%
  mutate(range = Uci - Lci,
         prop_range = range / N) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  mutate(model = fct_recode(model,
                            "Schumacher-Eschmeyer" = "SchumacherEschmeyer"),
         model = fct_relevel(model,
                             "Petersen",
                             after = 0)) %>%
  arrange(event_name, model) %>%
  # remove the Petersen and Schumacher-Eschmeyer estimators
  filter(! model %in% c("Petersen",
                        "Schumacher-Eschmeyer"))

```

```{r spring-abund}
total_cpue = cpue_df %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  group_by(species,
           event_name) %>%
  summarise(across(c(total_effort_min,
                     n_fish,
                     n_large_fish,
                     angler_days = crew_count),
                   sum),
            .groups = "drop") %>%
  mutate(catch_per_hr = n_fish / (total_effort_min / 60),
         catch_per_angler_day = n_fish / angler_days,
         catch_lg_per_angler_day = n_large_fish / angler_days)

spring_abund = all_est %>%
  filter(model == "Schnabel") %>%
  #filter(model == "Schnabel - Delayed Mixing") %>%
  select(-range, -prop_range) %>%
  inner_join(total_cpue,
             by = c("event_name", "species")) %>%
  select(species,
         model,
         model_event = event_name,
         N_F = N,
         Lci_F = Lci,
         Uci_F = Uci,
         cpue_F = catch_per_hr) %>%
  group_by(species, model) %>%
  summarise(across(c(N_F:cpue_F),
                   mean),
            .groups = "drop") %>%
    full_join(total_cpue %>%
              filter(str_detect(event_name, "Spring")) %>%
              select(species,
                     cpue_S = catch_per_hr),
            by = "species") %>%
  # calculate spring abundance estimate
  mutate(N_S = N_F * (cpue_S / cpue_F),
         Lci_S = Lci_F * (cpue_S / cpue_F),
         Uci_S = Uci_F * (cpue_S / cpue_F))

```

Two types of mark-recapture estimators were used to explore variation in abundance estimates: a single census and a multiple census estimator. For the single census estimator, we treated the first week of the survey as the mark event and the following week as the recapture event, pooling data within each week. Alternatively, the multiple census estimator treated each day as an event and used information about the total number of marked fish from all previous events to calculate the total abundance.

For the single census estimator, we used the Chapman-modified Lincoln-Peterson estimator:

$$
\hat{N} = \frac{(M + 1)(n + 1)}{(m + 1)} - 1
$$

where $M$ is the total number of fish marked and returned to the population during the first event (week), $n$ is the total number of fish caught in the second event (week), and $m$ is the number of marked fish caught during the second week. For a multiple census estimator, we used the Schnabel method:

$$
\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_i}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
$$

where $M$, $n$, and $m$ are indexed by each survey (day), $i$. With small numbers of recaptures, it may not be appropriate to calculate the standard error of the Schnabel estimator, but the 95% confidence intervals can be calculated using a Poisson approximation [@Krebs1999].<!--The Schnabel estimator is essentially the weighted average of a series of Chapman-modified Lincoln-Peterson estimators.-->

Because captures of new and marked fish were dependent on volunteer anglers, there was concern that sampling was biased to locations preferred by anglers. If marked fish were not recently released or mixed back into those preferred locales, marked fish may have been less susceptible to capture than unmarked fish. To account for this potential bias, we also used an adjusted Schnabel estimator, which assumed that marked and released fish during each survey day were not available for recapture for 48 hours. The adjusted Schnabel estimator, referred to as the "delayed mixing" Schnabel estimator, is given by:

$$
\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_{i-1}}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
$$

The mark-recapture estimators allowed us to estimate the abundance of Northern Pikeminnow during the fall 2019 and 2020 surveys. However, to estimate the abundance of Northern Pikeminnow during spring 2021, which was not a mark-recapture design, we used the ratio of total CPUE in the fall survey to the total CPUE in the spring survey, then multiplied that ratio by the mean abundance from fall surveys.

$$
\hat{N}_{spring} = \overline{N}_{fall} * \frac{CPUE_{spring}}{CPUE_{fall}}
$$

This approach assumes equal capture probabilities for fall and spring surveys. Fall and spring population estimates were converted into linear and areal densities by dividing by the length and area, respectively, of Deadwater Slough measured in GIS software and using drone-generated orthography.

All mark-recapture abundance estimators assumed: (1) the population is closed (no immigration, emigration, births or deaths during the survey period), (2) all fish have equal chance of being caught in subsequent survey days, (3) marking a fish does not affect its chances of recapture, (4) no loss of marks, and (5) marks are not missed or mistaken.

```{r psd}
s_size = 250        # stock size
q_size = large_size # quality size

psd_df = cmr_df %>%
  select(event_name,
         species,
         count,
         length,
         recap) %>%
  filter(recap != T)

fq = psd_df %>%
  filter(length >= q_size) %>%
  nrow()

fs = psd_df %>%
  filter(length >= s_size) %>%
  nrow()

psd = 100 * (fq/fs)

```

Anglers reported that Northern Pikeminnow in Deadwater Slough were large-bodied. To describe the population size structure, we calculated the proportional stock density (PSD) for Northern Pikeminnow in Deadwater Slough. PSD is a ratio, expressed as a percentage, between the number of "quality-sized" (larger) individuals and "stock-sized" individuals:

$$
PSD_{i} = 100 * \frac{FQ_{i}}{FS_{i}}
$$

where $FQ_{i}$ is the number of fish $\ge$ quality-length for species $i$, and $FS_{i}$ is the number of fish $\ge$ stock-length for species $i$. For Northern Pikeminnow in Deadwater Slough, we used `r prettyNum(q_size, digits = 0)` mm TL for quality-length and `r prettyNum(s_size, digits = 0)` mm TL for stock-length [@Winther2020].

## Stomach Contents

Gastric lavage [@Foster1977] was used to examine the stomach contents of Northern Pikeminnow for the presence of juvenile Chinook Salmon and other fishes occupying Deadwater Slough. Immediately following lavage, stomach contents of individuals were preserved with 99% isopropyl alcohol in whirl-paks  to be analyzed in a controlled environment. For each stomach sample, wet weight (grams) was recorded for the total stomach contents, including all non-fish items (e.g., macroinvertebrates, organic matter), as well as the portion consisting of fish. Fish and fish remnants were identified to the lowest taxonomic unit using diagnostic bones [@Hansel1988; @Frost2000] or were categorized as unknown. Approximately 5% of Northern Pikeminnow were euthanized for dissection (n = 75) after gastric lavage to validate the efficacy of the methodology.

```{r gastric-data}
gastric_df = raw_df %>%
  select(event_name, 
         date, 
         species,
         method,
         length,
         lavage_id,
         stomach_contents,
         stomach_contents_weight,
         fish_content_weight,
         comments) %>%
  #mutate(bin_50 = lencat(length, w = 50)) %>%
  filter(species == "Northern Pikeminnow",
         method == "Angling") %>%
  mutate(non_fish_wt = stomach_contents_weight - fish_content_weight) %>%
  mutate(lavaged = case_when(
    is.na(lavage_id) & is.na(stomach_contents) ~ FALSE,
    stomach_contents == "Did Not Lavage" ~ FALSE,
    stomach_contents == "Lost sample CG27-001" ~ FALSE,
    TRUE ~ TRUE
  )) %>%
  mutate(had_contents = case_when(
    lavaged == FALSE ~ FALSE,
    is.na(stomach_contents) ~ FALSE,
    stomach_contents == "Empty" ~ FALSE,
    TRUE ~ TRUE
  )) %>%
  mutate(had_fish_contents = case_when(
    fish_content_weight > 0 ~ TRUE,
    stomach_contents == "MacroInvert-other/Detritus/Fish-Unk" ~ TRUE,
    stomach_contents == "Fish-Unk (fish bone/cartilage)" ~ TRUE,
    stomach_contents == "MacroInvert-other/Mountain Whitefish" ~ TRUE,
    stomach_contents == "MacroInvert-other/Algae/Detritus/Fish-Unk" ~ TRUE,
    stomach_contents == "Fish -Shiner" ~ TRUE,
    stomach_contents == "Detritus/Fish-Unk" ~ TRUE,
    stomach_contents == "Fish Unk" ~ TRUE,
    stomach_contents == "Detritus/Fish -Shiner" ~ TRUE,
    stomach_contents == "MacroInvert-other/Fish-Unk" ~ TRUE,
    stomach_contents == "MacroInvert-other/Detritus/Fish -Shiner" ~ TRUE,
    stomach_contents == "Fish-Unk" ~ TRUE,
    stomach_contents == "MacroInvert-Other/Detritus/Fish-Unk" ~ TRUE,
    stomach_contents == "MacroInvert-Other/Detritus/Fish-Unk/Other"  ~ TRUE,
    stomach_contents == "Aquatic Veg/Fish-Sucker"  ~ TRUE,
    stomach_contents == "MacroInvert-Worm/Fish-RedsideShiner"  ~ TRUE,
    stomach_contents == "Aquatic Veg/Detritus/Fish-RedsideShiner"  ~ TRUE,
    stomach_contents == "MacroInvert-Other/Fish-Unk" ~ TRUE,
    stomach_contents == "Fish-Sucker/Other" ~ TRUE,
    stomach_contents == "Fish-Unk/Other" ~ TRUE,
    stomach_contents == "MacroInvert-Other/Aquatic Veg/Detritus/Fish-Sucker" ~ TRUE,
    stomach_contents == "Fish-RedsideShiner" ~ TRUE,
    stomach_contents == "Fish-Chinook/Other" ~ TRUE,
    stomach_contents == "Detritus/Fish-RedsideShiner/Other" ~ TRUE,
    stomach_contents == "MacroInvert-Other/Aquatic Veg/Detritus/Fish-RedsideShiner" ~ TRUE,
    stomach_contents == "Detritus/Fish-Sucker/Other" ~ TRUE,
    stomach_contents == "Invertebrates;FishParts"  ~ TRUE,
    stomach_contents == "Invertebrates;Detritus;FishParts;IntestinalWorms"  ~ TRUE,
    stomach_contents == "FishParts"  ~ TRUE,
    stomach_contents == "FishParts;IntestinalWorms"  ~ TRUE,
    TRUE ~ FALSE
  )) %>%
  mutate(had_chinook = case_when(
    stomach_contents == "Fish-Chinook/Other" ~ TRUE,
    TRUE ~ FALSE
  )) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " ")))

gastric_tbl = gastric_df %>%
  group_by(event_name) %>%
  summarise(captured = n(),
            lavaged = sum(lavaged),
            stmch_contents = sum(had_contents),
            fish_contents = sum(had_fish_contents)) %>%
  adorn_totals("row")

stmch_df = gastric_df %>%
  select(event_name,
         species,
         date,
         length,
         stomach_contents_weight,
         fish_content_weight,
         non_fish_wt) %>%
  filter(stomach_contents_weight > 0) %>%
  mutate(p_fish = fish_content_weight / stomach_contents_weight)

```

## Fish Consumption Potential

```{r shoup-temp-data, eval = F}
shoup_temp_df = read_tsv(here("analysis/data/raw_data/USGS_13307000_Salmon_River_Nr_Shoup.txt"),
                   comment = "#") %>%
  rename(discharge_cfs = `47768_00060`,
         discharge_cfs_cd = `47768_00060_cd`,
         gage_ht_ft = `47769_00065`,
         gage_ht_ft_cd = `47769_00065_cd`,
         water_temp_c = `47773_00010`,
         water_temp_c_cd = `47773_00010_cd`) %>%
  select(datetime, water_temp_c) %>%
  mutate(month = month(datetime),
         day = day(datetime)) %>%
  group_by(month, 
           day) %>%
  summarize(mean_water_temp = mean(water_temp_c, na.rm = T))

```

```{r length-weight}
# average size of fall and spring pikeminnow
avg_fall_NPM_TL = cmr_df %>%
  filter(grepl('Fall', event_name)) %>%
  select(length) %>%
  pull() %>%
  mean(na.rm = T)

avg_spring_NPM_TL = cmr_df %>%
  filter(grepl('Spring', event_name)) %>%
  select(length) %>%
  pull() %>%
  mean(na.rm = T)

# conversions from Fishbase.org
#TL = 1.052 * FL
#TL = 1.087 * FL (12-30cm)
#TL = 1.114 * FL (5-12cm) don't use

# use mean of coefficients
# TL = 1.0695 * FL or
# FL = 0.9350 * TL

# convert to FL
avg_fall_NPM_FL = avg_fall_NPM_TL * 0.9350
avg_spring_NPM_FL = avg_spring_NPM_TL * 0.9350

# convert FL to g using Parker (1995) equation
avg_fall_NPM_g = 10^(-4.886 + (2.986*log10(avg_fall_NPM_FL)))
avg_spring_NPM_g = 10^(-4.886 + (2.986*log10(avg_spring_NPM_FL)))

# average size of DSR and NRR juveniles
avg_DSR_g = 10.3
avg_NRR_g = 10.9

```

```{r read-bio-outputs, message=FALSE, warning=FALSE, include=FALSE}
fmt_bio_out = function(x) {
  x = x %>%
    clean_names() %>%
    mutate(date = as.Date(date, "%m/%d")) %>%
    mutate(cum_cons_fish_g = cumsum(cons_fish_g)) %>% 
    mutate(month = month(date))
  return(x)

}

# fall, fish 30% - 90%
fall_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I70_F30.csv")) %>% fmt_bio_out()
fall_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I60_F40.csv")) %>% fmt_bio_out()
fall_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I50_F50.csv")) %>% fmt_bio_out()
fall_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I40_F60.csv")) %>% fmt_bio_out()
fall_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I30_F70.csv")) %>% fmt_bio_out()
fall_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I20_F80.csv")) %>% fmt_bio_out()
fall_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I10_F90.csv")) %>% fmt_bio_out()

# spring, fish 30% - 90%
spring_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I70_F30.csv")) %>% fmt_bio_out()
spring_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I60_F40.csv")) %>% fmt_bio_out()
spring_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I50_F50.csv")) %>% fmt_bio_out()
spring_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I40_F60.csv")) %>% fmt_bio_out()
spring_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I30_F70.csv")) %>% fmt_bio_out()
spring_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I20_F80.csv")) %>% fmt_bio_out()
spring_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I10_F90.csv")) %>% fmt_bio_out()

# compile fall results
fall_results = fall_f30 %>%
  select(date, temperature_c, cum_cons_fish_g) %>%
  rename(f30_g = cum_cons_fish_g) %>%
  mutate(f40_g = fall_f40$cum_cons_fish_g,
         f50_g = fall_f50$cum_cons_fish_g,
         f60_g = fall_f60$cum_cons_fish_g,
         f70_g = fall_f70$cum_cons_fish_g,
         f80_g = fall_f80$cum_cons_fish_g,
         f90_g = fall_f90$cum_cons_fish_g) %>%
  mutate(season = "Fall") #%>%
  # mutate( across(contains('_g'),
  #                .fns = list(n = ~./avg_DSR_g),
  #                .names = "{col}_{fn}") )

# compile spring results
spring_results = spring_f30 %>%
  select(date, temperature_c, cum_cons_fish_g) %>%
  rename(f30_g = cum_cons_fish_g) %>%
  mutate(f40_g = spring_f40$cum_cons_fish_g,
         f50_g = spring_f50$cum_cons_fish_g,
         f60_g = spring_f60$cum_cons_fish_g,
         f70_g = spring_f70$cum_cons_fish_g,
         f80_g = spring_f80$cum_cons_fish_g,
         f90_g = spring_f90$cum_cons_fish_g) %>%
  mutate(season = "Spring") # %>%
  # mutate( across(contains('_g'),
  #                .fns = list(n = ~./avg_NRR_g),
  #                .names = "{col}_{fn}") )

bio_results = rbind(fall_results, spring_results)
bio_g_summ = bio_results %>%
  select(season, f30_g:f90_g) %>%
  pivot_longer(f30_g:f90_g) %>%
  group_by(season, name) %>%
  summarize(cumm_g = max(value))

rm(fall_results,
   fall_f30, 
   fall_f40, 
   fall_f50, 
   fall_f60, 
   fall_f70, 
   fall_f80, 
   fall_f90,
   spring_results,
   spring_f30, 
   spring_f40, 
   spring_f50, 
   spring_f60, 
   spring_f70, 
   spring_f80, 
   spring_f90)

```

To estimate the total consumption potential of Northern Pikeminnow in Deadwater Slough during the peaks of fall DSR and spring NRR emigrations, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 applied in R statistical software [@RCoreTeam2021]. The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures. Separate models were run for the fall period (September 15 - November 30) and the spring period (March 1 - May 31) to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG.

Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 3,000 J/g for invertebrates and 21,500 J/g for juvenile Chinook Salmon [@Moss2016]. Because we were unable to differentiate juvenile Chinook Salmon from other fish prey, we assume all fish prey have the same energy densities as juvenile Chinook Salmon. The average TL of Northern Pikeminnow caught in Deadwater Slough during our study was converted to fork length (FL) and then to weight (grams) using a weight-length formula from @Parker1995.

$$
log_{10} W_s = -4.886 + 2.986[log_{10}(FL)]
$$

This resulted in average fall and spring starting weights of `r round(avg_fall_NPM_g, 1)` and `r round(avg_spring_NPM_g, 1)` g for Northern Pikeminnow with average TLs of `r round(avg_fall_NPM_TL, 1)` and `r round(avg_spring_NPM_TL, 1)` mm, respectively. We assumed no growth in individual Northern Pikeminnow, resulting in equal start and end weights. Mean daily water temperatures were summarized from 15-minute interval temperature readings between March 3, 2013 and June 14, 2021 from USGS gage station 13307000, approximately 22 rkm downstream of Deadwater Slough.

The proportion of fish in the Northern Pikeminnow's diet relative to non-fish prey items is unknown in Deadwater Slough. Therefore, we conducted separate model runs that assume different proportions of fish in the diet ranging from 30 - 90%, in 10% increments. These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey. Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (Table \@ref(tab:bio-tbl)). To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.

## Impacts to Chinook Salmon Populations

The proportion of Chinook Salmon prey relative to other fish prey in the Northern Pikeminnow diet is also unknown. However, there is some evidence that juvenile salmonids are by far the most consumed fish prey [@Zimmerman1999; @Shively1996]. Moreover, it is estimated that Chinook Salmon make up 64.2%, 29.3%, and 49.3% of the fish prey consumed by Northern Pikeminnow in the Columbia River below Bonneville Dam, in Columbia River reservoirs, and in the lower Snake River, respectively [@Zimmerman1999]. Because the bioenergetics model does not differentiate juvenile Chinook Salmon from other fish prey items, we performed a sensitivity analysis to assess potential impacts of Northern Pikeminnow predation on local Chinook Salmon populations. Using values similar to @Zimmerman1999, we modeled three diet scenarios where Chinook Salmon comprised 30%, 50%, and 65% of the fish prey, by weight, in the Northern Pikeminnow diet during the fall DSR and spring NRR emigration periods. In this analysis, we assume 60% of the Northern Pikeminnow's total diet is comprised of fish, representing the median scenario in our range of modeled diets.

The estimate of total juvenile Chinook Salmon biomass consumed was then converted to the total estimated number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG. The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods. Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns. We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation was reduced or eliminated in the Deadwater Slough. This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by the median Granite-to-Granite SAR of 0.00614 (sd = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.

```{r adult-impacts}
# pull in SAR numbers
sar = read_csv(here("analysis/data/raw_data", 
                    "WSSCh_sar_4See.csv")) %>%
  summarise(across(c(medsar, sdsar),
                   median)) %>%
  summarise(across(c(medsar, sdsar),
               ~ . / 100))

# summarize potential Chinook eaten for each time period by one pikeminnow
impact_df = bio_results %>%
  group_by(season) %>%
  summarize(across(ends_with("_g"), max)) %>%
  pivot_longer(cols = ends_with("_g"),
               names_to = "perc_diet",
               values_to = "g_fish_cons") %>%
  mutate(across(perc_diet,
                str_extract,
                "[:digit:]+"),
         across(perc_diet,
                as.integer)) %>%
  #filter(perc_diet == 60) %>%
  crossing(p_chnk = c(.30, .50, .65)) %>%
  mutate(g_chnk = g_fish_cons * p_chnk) %>%
  mutate(n_chnk = ifelse(season == "Fall",
                         g_chnk / avg_DSR_g,
                         g_chnk / avg_NRR_g)) %>%
  left_join(spring_abund %>%
            select(N_F:Uci_F,
                   N_S:Uci_S) %>%
            pivot_longer(everything()) %>%
            mutate(season = if_else(str_detect(name, "_F$"),
                                    "Fall",
                                    "Spring")) %>%
            mutate(across(name,
                          str_remove,
                          "_F$"),
                    across(name,
                           str_remove,
                           "_S$")) %>%
            pivot_wider(names_from = name,
                        values_from = value)) %>%
  mutate(juv_N = N * n_chnk,
         juv_Lci = Lci * n_chnk,
         juv_Uci = Uci * n_chnk) %>%
  bind_cols(sar) %>%
  mutate(adult_N = juv_N * medsar,
         adult_Lci = juv_Lci * qnorm(0.025, medsar, sdsar),
         adult_Uci = juv_Uci * qnorm(0.975, medsar, sdsar))

# sum fall and spring impacts
total_impacts = impact_df %>%
  group_by(perc_diet, p_chnk) %>%
  summarise(across(c(ends_with("_N"),
                     ends_with("_Lci"),
                     ends_with("_Uci")),
                   sum),
            .groups = "drop") %>%
  select(perc_diet,
         p_chnk,
         juv_N,
         juv_Lci,
         juv_Uci,
         adult_N,
         adult_Lci,
         adult_Uci)

```

All data and code for the analyses presented here can be found in a GitHub repository at ***zenodo DOI*** here.

# Results

## Northern Pikeminnow Demographics

Using hook-and-line angling, we caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 surveys, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish), big.mark = ",")` fish. The CPUE was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021. During the mark-recapture events, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019, including `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals and `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020, including `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals. Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)). For both surveys, the multiple census estimates were generally larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)). Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate. Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`). Using the delayed-mixing Schnabel estimator, the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`). For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough. Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.

The lengths of Northern Pikeminnow captured ranged from `r min(cmr_df$length, na.rm = T)` to `r max(cmr_df$length, na.rm = T)` mm TL with an average of `r round(mean(cmr_df$length, na.rm = T))` mm (Figure \@ref(fig:length-freq-p)). The PSD for Northern Pikeminnow in Deadwater Slough across all three efforts was `r prettyNum(psd, digits = 1)`%.

## Stomach Contents

We found that gastric lavage successfully removed food items in all dissected individuals, indicating the method was effective. Of the `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow sampled using gastric lavage, we found contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs. Northern Pikeminnow captured during spring 2021 had a higher proportion of fish content in their diet compared to individuals captured in the fall surveys (Table \@ref(tab:gastric-tab)). The mean wet weight of total contents for an individual was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; sd = `r round(sd(stmch_df$stomach_contents_weight), 2)` g). Overall, fish or fish remnants made up `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of all stomach contents examined.

## Fish Consumption Potential

```{r tons-consumption}
tot_g_DSR = (bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"] * mean(all_est$N[all_est$model == "Schnabel"])) / 1000000
tot_g_NRR = (bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"] * round(spring_abund$N_S)) / 1000000

```

During the fall DSR emigration, we estimate an average-size Northern Pikeminnow (`r round(avg_fall_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f60_g"]` g of fish to maintain their body size, assuming 60% of their diet consists of fish prey (Figure \@ref(fig:bio-p)). Fall consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f30_g"]` g for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Fall" & bio_g_summ$name == "f90_g"]` g for a diet of 90% fish. During the spring NRR emigration, we estimated an average-size Northern Pikeminnow (`r round(avg_spring_NPM_TL, 1)` TL mm) to consume `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f60_g"]` g of fish, assuming 60% of their diet is fish prey (Figure \@ref(fig:bio-p)). Spring consumption ranged from `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f30_g"]` for a diet of 30% fish to `r bio_g_summ$cumm_g[bio_g_summ$season == "Spring" & bio_g_summ$name == "f90_g"]` for a diet of 90% fish. Given estimated mean Northern Pikeminnow population sizes of `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` from fall estimates and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` for spring 2021, and assuming 60% of their diet is fish, we estimate that Northern Pikeminnow population can consume `r round(tot_g_DSR, 1)` metric tons of fish during the fall DSR emigration and an additional `r round(tot_g_NRR, 1)` metric tons during the spring NRR emigration in Deadwater Slough.

## Impacts to Chinook Salmon Populations

Using the total fish consumed by the Northern Pikeminnow population (summing fall and spring surveys), we produced a sensitivity analysis for three Northern Pikeminnow diet scenarios where juvenile Chinook Salmon make up 30%, 50%, and 65% of total fish prey consumed (Figure \@ref(fig:juv-con-p)). Assuming that 50% of the total fish consumed are juvenile Chinook salmon, we estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$juv_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$juv_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) juvenile Chinook Salmon will be consumed. Using the median Granite-to-Granite SAR, we estimated the juvenile consumption would is equivalent to `r prettyNum(round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` (95% CI: `r prettyNum(round(total_impacts$adult_Lci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` - `r prettyNum(round(total_impacts$adult_Uci[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")`) adults (Figure \@ref(fig:adult-imp-p)).

# Discussion

## Mark-Recapture Models

```{r john-day}
john_day_res_abund = 85316
john_day_rkms = 123
john_day_dens = john_day_res_abund / john_day_rkms

low_columbia_fish_km = 2814

```

All mark-recapture abundance estimators suggest a large population of Northern Pikeminnow occupy Deadwater Slough. The mean population size of Northern Pikeminnow was estimated to be `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` during the fall DSR emigration and `r prettyNum(round(spring_abund$N_S), big.mark = ",")` during the spring NRR emigration. Those estimates translate to linear densities (fish/rkm) that are `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / john_day_dens)` and `r round((spring_abund$N_S / ddw_rkms) / john_day_dens)` times greater than densities reported by @Beamesderfer1991 for the John Day Reservoir and approximately `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_rkms) / low_columbia_fish_km)` and `r round((spring_abund$N_S / ddw_rkms) / low_columbia_fish_km)` times greater than projections for the lower Columbia River [@Beamesderfer1996]. @Beamesderfer1991 acknowledge that their methods were unsuitable for sampling offshore, noting that water velocity, depth, irregular bottom contours, and barge traffic made sampling offshore ineffective; therefore it is possible that their estimates may have been low. Conversely, at Deadwater Slough the maximum depth was approximately 6 m with a relatively homogenous bottom contour, making angling an effective method throughout the entire reach. Given the size (volume) of the Salmon River at the slough relative to the Columbia River, the estimated Northern Pikeminnow densities for Deadwater Slough are remarkable. Further, our estimate densities of Northern Pikeminnow suggest that slow water reaches outside of the reservoir complexes on the Snake and Columbia rivers may support exceptionally high predator densities, consistent with findings of @Harnish2014, @Gray2001, and @Zimmerman1999. Given the dearth of current data available in the Columbia River basin, it is unclear how Northern Pikeminnow abundance estimates in this study compare to elsewhere in the Columbia River Basin; contemporary information on piscine predator abundances may be needed [@Widener2021].

Several assumptions in our estimators may have influenced the magnitude of our abundance results. First, all models assumed a closed population, however, it is possible that immigration and emigration occurred. Nevertheless, our estimators still provide unbiased estimates of abundance assuming the immigration and emigration rates are equal between marked and unmarked fish. Given the size of the sampling area, and the short duration of our surveys, that assumption is likely met. <!--We also caught one individual that was confirmed tagged in Deadwater Slough in a previous year suggesting that at least some individuals remain in the reach or at least migrate to Deadwater Slough multiple times, annually.-->The assumption of a closed population can also affect results if mortality occurs for some marked fish released back to the population. During field processing, no mortalities or injury post release was observed; however, the potential for mortality amongst marked and released fish cannot be discounted. Next, our estimates assumed that capture and recapture events are random samples of the population. Angling methods may have a size selection bias, thereby limiting our abundance estimates to a size range susceptible to angling, resulting in a conservative estimate of the total population. We also question the validity of our assumption of equal catchability of individual fish between sampling events. During our study, we recaptured two individual fish during multiple days within a survey indicating that marked fish were still susceptible to angling.<!--We are also comfortable with the assumptions of no loss of marks or no mistaken or missed marks given that marks were physical (hole punches in fins) and that each fish was examined on a measuring board to collect lengths.--> However, we suspect that marked and released fish may not have mixed thoroughly back into the population prior to recapture events. Although we attempted to release marked fish evenly throughout Deadwater Slough, we anecdotally observed anglers congregating in particular areas of the slough to socialize or exploit the "good" fishing locales. Violations of the assumption of equal catchability between marked and unmarked fish may lead to an overestimate of abundance of Northern Pikeminnow in our study if marked fish have lower catchability. This bias will be proportional to the difference in catchability of marked and unmarked fish. For example, if marked fish were only 70% as likely to be caught as unmarked fish, the true abundance is closer to 70% of our abundance estimate.

The spring abundance estimate was smaller than fall abundance estimates. Although our approach assumed equal capture probabilities between the fall and spring surveys, we believe higher spring flows may result in lower capture probability. Therefore, we believe our spring abundance estimate to be conservative.

In addition to the large population, the observed PSD (`r round(psd)`%) in this study was large compared to observations of 41% below Bonneville Dam and 18% in Bonneville Reservoir [@Winther2020]. This suggests that a larger fraction of Northern Pikeminnow in Deadwater Slough are of a quality size relative to populations reported elsewhere in the Columbia River. Notably, the state catch-and-release record Northern Pikeminnow, measuring `r max(gastric_df$length, na.rm = T)` mm TL, was caught in Deadwater Slough during the fall 2020 survey.

## Stomach Contents

We performed gastric lavage on nearly all Northern Pikeminnow collected during this study. Observed prey species included juvenile Chinook Salmon as well as Redside Shiner *Richardsonius balteatus*, Largescale Sucker *Catostomus macrocheilus*, sculpin *Cottus* spp., and Mountain Whitefish *Prosopium williamsoni*, though decomposition from digestion rendered most fish contents unidentifiable. Although our approach provided useful information on prey species, we have concerns about using angling and gastric lavage to quantify diet composition in Northern Pikeminnow. First, diets of angled fish can differ from those collected by other methods such as electrofishing [@Hodgson1988]. Further, angled fish are more likely to have empty stomachs [@Jurajda2016], perhaps because they are hungry and actively searching for food. Second, unlike other predators such as Smallmouth Bass and Walleye, that are commonly sampled using gastric lavage [@Kamler2001], Northern Pikeminnow are cyprinids and therefore lack a true stomach. Consequently, performing gastric lavage on cyprinid species can be ineffective for accurately quantifying diet composition [@Hartleb1995], or alternately, requires a modification in methodology to flush food items out the vent of the fish [@Wasowicz1994]. Therefore, we consider our diet composition estimates and subsequent consumption estimates to be conservative.

## Bioenergetics

Our approach assumed that all available prey consumed by Northern Pikeminnow have energy densities equal to juvenile Chinook Salmon. Generalizing energy densities using a single species is a common approach [@Petersen1999], however, we recognize that other salmonids including juvenile steelhead and Sockeye Salmon are likely also available prey in Deadwater Slough, especially during the spring migration. Additionally, hatchery smolt releases (Chinook Salmon, steelhead, and Sockeye Salmon) are also prevalent in the mainstem Salmon River during the spring migration. To address the prevalence of other fish in the Northern Pikeminnow's diet, we considered scenarios where DSR and NRR Chinook Salmon represent only 30%, 50%, and 65% of total fish prey consumed. These values may be considered conservative during the emigration periods when Chinook Salmon are abundant and the Northern Pikeminnow diet shifts almost entirely to piscivory, presumably on juvenile Chinook Salmon [@Shively1996; @Poe1991]. This is likely the case during the fall when DSR Chinook Salmon are the dominant prey species available in Deadwater Slough as 1) no (or at least minimal) hatchery releases are present in the river and 2) fall emigrations of steelhead and Sockeye Salmon are less prominent that in Chinook Salmon. During spring when natural-origin steelhead and Sockeye Salmon are also actively emigrating from the Upper Salmon River and hatchery-origin releases of all three species (Chinook Salmon, steelhead, Sockeye Salmon) are present in the river, <50% Chinook Salmon in the Pikeminnow's diet may be more likely. Future work to quantify and identify the juvenile salmonid in Deadwater Slough during their seasonal migration would be useful to validate the diet composition assumptions used in our model.

## Avian Predation

Although not formally assessed in this study, avian predators are another potential source of mortality for juvenile Chinook Salmon and other ESA-listed salmonids in the Deadwater Slough. The Deadwater Slough is recognized as an important bird watching and nesting area due to the associated upstream braided channels and riparian and backwater habitats [(Deadwater Slough - Audubon Important Bird Areas)](https://www.audubon.org/important-bird-areas/deadwater-slough#). Several piscivorous bird species have been documented using Deadwater Slough including the Common *Mergus merganser* and Hooded *Lophodytes cucullatus* mergansers, the Great Blue Heron *Ardea herodias*, the Double-crested Cormorant *Phalacrocorax auritus*, and the Belted Kingfisher *Megaceryle alcyon* [@eBird2021]. During the initial sampling effort in fall 2019, a two-person crew walked the entire reach including surrounding and upstream areas scanning for passive integrated transponder (PIT) tags that are implanted into thousands of natural- and hatchery-origin salmonids in the Upper Salmon River, annually. During that informal survey, 10 PIT tags were recovered near active bird nests and in an upstream anastomizing reach where herons and eagles are prevalent, suggesting that mortality may have been a result of avian predation. The PIT tag histories in [PTAGIS](https://www.ptagis.org/) indicate these tags were implanted into a combination of juvenile Chinook Salmon (3), Sockeye Salmon (4), and steelhead (3). Two of the Chinook Salmon and one steelhead were natural-origin, the remaining were of hatchery-origin. Avian predation contributes a major component of the total mortality for yearling Chinook Salmon in some locations in the lower Snake and Columbia rivers, particularly at hydroelectric dams and within reservoirs [@Evans2012; -@Evans2016]; however, we did not observe large colonies of piscivorous birds within the study area. Although we did observe Belted Kingfishers, mergansers, and Great Blue Herons at Deadwater Slough; Great Blue Herons were especially prevalent in the anastomizing reach immediately upriver of the slack water reach. Additionally, there is documentation of individual Double-crested Cormorants [@eBird2021] at the Deadwater Slough; however, the site is not within their breeding range, rather, it is part of a migration corridor. Given the current avian species known to occupy Deadwater Slough, it is unlikely that avian predation on juvenile salmonids is comparable to elsewhere in the Columbia River basin with large piscivorous bird colonies. Nevertheless, we hypothesize that the reservoir-like conditions and associated habitat features at the Deadwater Slough may increase the probability of avian predation on juvenile Chinook Salmon and other ESA-listed salmonids from the many piscivorous birds known to use the site. Future studies of predation would benefit from estimates from piscivorous avian predators.

## Potential Impacts to Salmon Populations

Recovery of Chinook Salmon in the Upper Salmon MPG is desired to support local fisheries and economies. At least five of the eight populations must meet criteria set forth by @McElhany2000 and the @ICTRT2007 for the MPG to be considered viable and for recovery of the Snake River Evolutionary Significant Unit. We estimate that `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50]), big.mark = ",")` juvenile Chinook Salmon may be consumed by Northern Pikeminnow assuming that 60% of their diet is fish and 50% of that is Chinook Salmon. Considering that diet scenarios used in our study are likely conservative assumptions, especially during the fall DSR emigration, the estimated number of Chinook Salmon consumed is substantial.

Spawner abundance is perhaps the most important metric considered in determining a population's viability. We estimated that consumption of juvenile Chinook Salmon by Northern Pikeminnow in the Deadwater Slough potentially reduces natural-origin returns to upriver populations by `r round(total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50])` adults, annually, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50] / 558) * 100)`% of the total mean adult escapement estimated<!--of 558--> for the Upper Salmon MPG in 2017 - 2019 [@Kinzer2020]. Even under the most conservative scenario where 30% of the Northern Pikeminnow diet is fish and 30% of that is Chinook Salmon we estimate the "adult equivalents" of the juveniles consumed to be `r round(total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30])`, which is `r round((total_impacts$adult_N[total_impacts$perc_diet == 30 & total_impacts$p_chnk == 0.30] / 558) * 100)`% of recent natural-origin adult escapements. In addition to juvenile Chinook Salmon, juvenile steelhead and Sockeye Salmon, including hatchery releases, are also likely prey items for Northern Pikeminnow in Deadwater Slough. Accordingly, we surmise that adult returns of steelhead and Sockeye Salmon are also affected by predation, including hatchery populations, which provide recreational fishing opportunities. Consequently, reducing predation mortality at Deadwater Slough could potentially benefit multiple upriver natural and hatchery populations, including other ESA-listed species. Because the Deadwater Slough is part of the migratory pathway for multiple species and populations of emigrating salmonids, the positive impact could be greater than individual tributary rehabilitation actions which typically benefit a single population.

## Management Implications

The Deadwater Slough is a favorable candidate for management or restoration actions to benefit the local Chinook Salmon populations. Two potential management actions could reduce predation at Deadwater Slough: 1) removing the Dump Creek alluvial fan and 2) a local Northern Pikeminnow bounty program to encourage harvest in Deadwater Slough aimed at reducing the predator population size. Removing the Dump Creek alluvial fan has the benefit of restoring the natural fluvial processes in the Salmon River that likely existed in the reach prior to the failure of the mining reservoir dam in Dump Creek. Restoring natural processes could reduce Northern Pikeminnow densities to levels similar to upstream and downstream reaches where higher survival and transition probabilities have been observed compared to Deadwater Slough [@Axel2015; @Ackerman2018; @Porter2019]. Managers ought also consider the feasibility and net benefit of restoring fluvial processes relative to potential losses for recreational fishing and bird watching opportunities.

Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (@Winther2020). A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues. Bounty programs could be conducted year-round or seasonally to coincide with peak juvenile outmigration(s). Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.

## Conclusions

The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow. Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkable density given the size of the area. Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon. We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than 24,000 juvenile Chinook salmon (the lower 95% CI of the most conservative scenario evaluated). It is more likely that this value is closer to 100,000, annually, creating a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon MPG.

\newpage

# Acknowledgements

The authors extend much appreciation to the many volunteers that assisted with field efforts including collaborators from the Bureau of Reclamation, Idaho Governor’s Office of Species Conservation, Idaho Department of Fish and Game, and the Lemhi Regional Land Trust, among others. Special thanks to Jared Barker, Brian Hamilton, Tulley Mackey, and Chelsea Welke for their help in the field. We further appreciate the administrative support and guidance from Caitlin Alcott and Inter-Fluve. This manuscript benefited from a review by Sean Gibbs. Funding for this study was provided by the Bureau of Reclamation, Pacific Northwest Regional Office (contract No. 140R1021F0018) with assistance from the Idaho Governor’s Office of Species Conservation.

\newpage

# Literature Cited

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

# Tables

<div id="tables"></div>

```{r bio-tbl}
bio_tbl = bio_g_summ %>%
  mutate(`Date Range` = season) %>%
  select(season,
         `Date Range`,
         name,
         cumm_g) %>%
  mutate(name = str_sub(name, 2, 3)) %>%
  rename(Season = season,
         `% Fish in Diet` = name,
         `g Consumed` = cumm_g) %>%
  mutate(`Date Range` = recode(`Date Range`,
                               Fall = "Sep 15 - Nov 30",
                               Spring = "Mar 1 - May 31")) %>%
  kable(booktabs = T,
        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the fall DSR and spring NRR Chinook Salmon emigrations. The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown. Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
  kable_styling()

bio_tbl

```

\newpage

```{r abund-data-lp}
LP_mod %>%
  select(event_name, species, data) %>%
  distinct() %>%
  select(-species) %>%
  unnest(data) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  rename(`Survey` = event_name) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of Northern Pikeminnow angling survey data for the single-census mark-recapture estimator. M is the total number of fish caught in the first sample (week), marked and returned to the population; n is the total number of fish caught in the second sample (week), and m is the number of previously marked fish caught in the second week.") %>%
  kable_styling()

```

\newpage

```{r abund-data-schn}
N_mods %>%
  select(event_name, species, data) %>%
  distinct() %>%
  select(-species) %>%
  unnest(data) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  select(event_name, date, M, n, m, u, R) %>%
  rename(`Survey` = event_name,
         Date = date) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of Northern Pikeminnow angling survey data for standard Schnabel estimator. M is the total number of fish caught in previous samples (days), marked and returned to the population; n is the total number of fish caught in the current sample (day); m is the number of marked fish caught in that sample; u is the number of unmarked fish caught, and R is the number of marked fish returned to the population. Dates of sampling are also provided.") %>%
  kable_styling()

```

\newpage

```{r abund-tab}
all_est %>%
  mutate(across(c(N, SE, Lci, Uci),
                ~ prettyNum(round(.), big.mark = ","))) %>%
  rowwise() %>%
  mutate(CI = paste(Lci, Uci,
                    sep = " - ")) %>%
  select(-c(species, range:prop_range),
         -Lci, -Uci) %>%
  rename(`Survey` = event_name,
         Estimator = model,
         `95% CI` = CI) %>%
         # `L CI` = Lci,
         # `U CI` = Uci) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Point estimates, standard errors, and 95% confidence intervals for each of the mark-recapture estimators and for the fall sampling efforts. Standard errors are not available for the multiple-census estimators.") %>%
  kable_styling()

```

\newpage

```{r gastric-tab}
gastric_tbl %>%
  rename(Survey = event_name,
         Captured = captured,
         Lavaged = lavaged,
         'Stomach Contents' = stmch_contents,
         'Fish Contents' = fish_contents) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of gastric lavage results including the total number of fish captured and gastric lavaged. For Northern Pikeminnow that were gastric lavaged, we show the number containing stomach contents, and of those, the number containing fish contents.") %>%
  kable_styling()

```

```{r adult-imp-tab, eval = F}
impact_tbl = total_impacts %>%
  select(perc_diet,
         p_chnk,
         N = juv_N,
         Lci = juv_Lci,
         Uci = juv_Uci) %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  mutate(Stage = "Juveniles") %>%
  bind_rows(total_impacts %>%
              select(perc_diet,
                     p_chnk,
                     N = adult_N,
                     Lci = adult_Lci,
                     Uci = adult_Uci) %>%
              mutate(p_chnk = p_chnk * 100) %>%
              rename(perc_chnk = p_chnk) %>%
              mutate(Stage = "Adult Equivalents")) %>%
  select(Stage,
         `% Chinook` = perc_chnk,
         N,
         `95% L CI` = Lci,
         `95% U CI` = Uci) %>% 
  mutate(across(where(is.numeric), round)) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Estimates of juvenile Chinook Salmon consumed and adult equivalents for scenario where 60% of Northern Pikeminnow diet is fish, the median of scenarios evaluated. Table shows scenarios where 30%, 50%, and 65% of total fish consumed is Chinook Salmon. Estimates and 95% confidence intervals are calculated using Northern Pikeminnow abundance estimates of 27,874 (95% CI: 14,244 - 59,388) for fall and 14,799 (95% CI: 7,562 - 31,530) for spring. 'Adult equivalents' are estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for upper Salmon River Chinook Salmon from McCann et al. (2019).") %>%
  kable_styling()

impact_tbl

```

\newpage

# Figures

<div id="figures"></div>

```{r map, out.width = "85%", fig.cap = "Map of the Deadwater Slough study area. The high-resolution orthoimage portion surrounding the Deadwater Slough was generated from aerial images taken by an unmanned aircraft system (drone). The red polygon indicates the reach characterized by unnaturally slow water velocities and a deepened channel. The location of the Dump Creek delta is indicated."}
knitr::include_graphics("../figures/deadwater_map_low_res.png")

```

\newpage

```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow in Deadwater Slough using different mark-recapture estimators and for the fall sampling events. Error bars indicate 95% confidence intervals."}

pos = position_dodge(width=0.3)
pm_est_labels <- c("Chapman","Schnabel","Schnabel \n [Delayed-mixing]")
all_est %>%
  mutate(across(c(N, Lci, Uci),
                ~ . / 1000)) %>%
  mutate(census = case_when(grepl("Chapman", model) ~ "single",
                            grepl("Petersen", model) ~ "single",
                            grepl("Schnabel", model) ~ "multiple",
                            grepl("Schumacher-Eschmeyer", model) ~ "multiple")) %>%

  ggplot(aes(x = model,
             y = N,
             #color = model
             )) +
  geom_errorbar(aes(ymin = Lci,ymax = Uci, group=event_name), position = pos, width=0) + 
  geom_point(aes(shape = event_name), position = pos, size = 4) +
    #facet_wrap(~ event_name) +
  coord_cartesian(ylim = c(0, 80)) +
  scale_shape_manual(values = c(21, 19), name="Survey") +
  #scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
  scale_x_discrete(labels=c(pm_est_labels))+
  labs(y = "Estimated Abundance (x 1,000)",
       x = "Estimator") +
  #scale_size(guide = "none")+
  theme(legend.position = c(.15, .85),legend.background = element_blank(),
        legend.key.size = unit(.5, "cm"),legend.box = 'horizontal',
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 11, color = "black"),
        axis.text.y = element_text(size = 11, color = "black"),
        strip.text.x = element_text(size = 11,color = "black"))
  #scale_color_brewer(palette = "Dark2")

```

\newpage

```{r length-freq-p, fig.cap = "Length frequency histogram of Northern Pikeminnow caught using hook-and-line angling during the study."}
length_freq_p = gastric_df %>%
  select(event_name,
         length) %>%
  mutate(season = event_name) %>%
  mutate(season = recode(season,
                         `Fall 2019` = "Fall",
                         `Fall 2020` = "Fall",
                         `Spring 2021` = "Spring"))

ggplot(length_freq_p, 
       aes(x = length, 
           fill = season)) +
  geom_histogram(breaks = seq(150, 650, by = 25),
                 col = "black",
                 alpha = 0.5) +
  scale_fill_manual(name = "Season",
                    values = c("gray", "black"),
                    labels = c("Fall", "Spring")) +
  labs(x = "Total Length (mm)",
       y = "Number of Northern Pikeminnow") +
  theme(axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        legend.position = c(.15, .85))

```

\newpage

```{r bio-p, fig.cap = "The cumulative fish (g) pre consumed by an individual Northern Pikeminnow during the peak DSR (fall) and NRR (spring) Chinook Salmon emigrations (top panel) and corresponding daily mean water temperatures (bottom panel). The black line is the median among diet scenarios (60% of the Northern Pikeminnow diet consists of fish) and the gray area shows the range between diet scenarios of 30 and 90% fish prey. Daily mean water temperatures were summarized from six years of data available from the Shoup gage station. "}
# spring_results$season = "Spring"
# fall_results$season = "Fall"
# bio_results = rbind(spring_results, fall_results)
bio_results$ymd = (ymd(bio_results$date))

pFish = bio_results %>%
  ggplot(aes(ymd, f60_g)) +
  geom_line(size = 1.1) +
  geom_ribbon(data = bio_results,
              aes(ymd,ymin = f30_g,
                  ymax = f90_g),
              color = "grey55",
              alpha = 0.4) +  theme_bw() +
  labs(y = "Fish consumed (g)",
       x = "", title = "") +
  facet_rep_wrap(~season,scales = "free_x",
                 repeat.tick.labels = F) +
  theme(strip.background = element_rect(color = "white",
                                        fill = "white"),
        strip.text.x = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        axis.title.y = element_text(vjust = 1,
                                    size = 12),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.margin = unit(c(-.1, .3, -.3, .2), "cm"))

pTemp = bio_results %>%
  ggplot(aes(ymd, temperature_c)) +
  geom_line(color = "navyblue",
            size = 1.15) +
  theme_bw() +
  labs(y = expression("Water temp"*~degree*C),
       x = "Month",
       title = "") +
  facet_rep_wrap(~season,
                 scales = "free_x",
                 repeat.tick.labels = F) +
  theme(strip.background = element_rect(color = "white",
                                        fill = "white"),
        strip.text.x = element_blank(),
        axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        axis.title.y = element_text(vjust = 0.1,
                                    size = 12),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.margin = unit(c(-.2, .3, .1, .1), "cm"))

p_list = list(pFish, pTemp)
n_col = floor(sqrt(length(p_list)))

bio_p = do.call("grid.arrange",
                c(p_list, ncol = n_col))

```

\newpage

```{r juv-con-p, fig.cap = "test"}
juv_cons_p = impact_df %>%
  select(season,
         perc_diet,
         p_chnk,
         N = juv_N,
         Lci = juv_Lci,
         Uci = juv_Uci) %>%
  group_by(perc_diet,
           p_chnk) %>%
  summarise(across(c("N", "Lci", "Uci"),sum),
            .groups = "drop") %>%
  mutate(Stage = "Juveniles") %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  ggplot(aes(x = perc_diet,
             y = N,
             group = as.factor(perc_chnk),
             color = as.factor(perc_chnk))) +
  geom_point(size = 2) +
  geom_line(size = 1.1) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
  scale_color_manual(name = "Percent Chinook", 
                     values=c("lightsteelblue","steelblue3","navyblue")) +
  theme(axis.text.y = element_text(size = 11,
                                   color = "black"),
        axis.text.x = element_text(size = 11,
                                    color = "black"),
        strip.text.y = element_text(size = 11,
                                    color = "black"),
        legend.position = c(.19, .89),
        legend.background = element_blank(),
        legend.key.size = unit(.56, "cm"),
        legend.box = 'horizontal') +
  labs(y = "Estimated Juvenile Chinook Consumed", 
       x = "Percent Fish in Pikeminnow Diet")

juv_cons_p

```

\newpage

```{r adult-imp-p, fig.cap = "temp"}
#adult impacts
adult_imp_p = total_impacts %>%
  filter(perc_diet == 60) %>%
  select(p_chnk,
         adult_N,
         adult_Lci,
         adult_Uci) %>%
  mutate(perc_chnk = p_chnk * 100) %>%
  select(-p_chnk) %>%
  ggplot(aes(x = as.factor(perc_chnk),
             y = adult_N)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = adult_Lci,
                    ymax = adult_Uci),
                width = 0) +
  labs(y = "Estimated Adult Chinook",
       x = "Percent Chinook in Diet") +
  theme(axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        legend.position = c(.2, .8),
        legend.box = "horizontal")

adult_imp_p

```

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()

```

\newpage

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())

```